Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-01-17 paper/manuscript.tex
Section 7 reviewed.

    \caption{\textbf{Thermodynamic cycle required for an aboslute free energy calculation}}

@@

For solvation free energy calculations, handling the standard state is typically straightforward, and treating it correctly simply means ensuring that the non-interacting solute is taken to the same (or equivalent) final reference state in both environments, e.g. that the transformation involves a 1M to 1M equivalent transfer free energy (where the non-interacting solute still occupies essentially the same volume as the solute in the interacting system).

@@

In practice, this has implications for how the calculations are done, as the reference concentration must enter the thermodynamic cycle employed.

@@

\todo[inline, color={blue!20}]{@author: Check references}
In practice, a variety of types of restraints are common, from simple harmonic distance restraints between the ligand and the protein~\cite{mobley2006use}, to flat-bottom restraints which work similarly but only exert a force if the ligand leaves a specific region~\cite{}.
Alternatively, a set of restraints proposed by Boresch have also commonly been employed, where all six rigid-body degrees of freedom governing the orientation of the ligand relative to the receptor are restrained~\cite{leitgeb2005alchemical}.

@@

For example, with more involved restraints, sampling at intermediate lambda values will usually not need to be as extensive but more computational effort must go to computing the restraining free energy.
Additionally, such restraints would typically keep the ligand from exploring alternative binding modes, which may be undesirable with Hamiltonian lambda exchange or expanded ensemble techniques where allowing the ligand to exchange binding modes when it is non-interacting could provide sampling benefits~\cite{wang2013identifying}.
Concretely, flat bottom restraints might allow a ligand to explore multiple binding sites, harmonic restraints multiple binding modes within a site, and Boresch restraints a single binding mode within a single site.
See additional discussion of the possibility of multiple binding modes below~\ref{sec:multiple_binding_modes}.

@@

If the ligand is a tight binder\todo[inline, color={red!40}]{AR: Should we give an order of magnitude to define a tight and a weak binder?}, it is usually safe to employ a restraint that allows some of the unbound configurations to be sampled since they generally contribute negligibly to the partition function (i.e., they have a relatively small Boltzmann weight) as long as the sampled volume is not so large that their cumulative contribution becomes significant.

@@

Many relative free energy planning tools have been set up to avoid changing the net charge of the systems considered, including LOMAP~\cite{} and Schr\"{o}dinger's FEP+~\cite{chen2018accurate}. Absolute free energy calculations can also potentially avoid changing the charge of the system by making a charge perturbation of equal and opposite sign elsewhere in the system; for example, as a charged ligand is removed, a charged counterion of opposite sign could also be removed, or one of the same sign could be inserted. This is sometimes referred to as an "alchemical ion" approach for dealing with the needed charge change, and is also employed by the Yank free energy package~\cite{wang2013identifying}.
Charge corrections have also been explored, and are potentially a viable solution to this problem~\cite{} where artifacts introduced by finite-size effects are corrected numerically.

@@

Some choices are particularly crucial -- for example, transformations involving insertions or deletions of atoms should employ a path soft-core potential path for Lennard-Jones or other interactions with repulsive interactions that go to infinite energy at small radius~\cite{}. 

@@

Another way of stating this is that intermediate states should sample molecular configurations that are as similar as possible to their neighboring states. The more similar the configurations are between intermediate states, the lower the statistical uncertainty is in the estimate of free energy between intervals.  This can be proven directly from the BAR and MBAR formulas~\cite{bennett1976efficient,klimovich2015guidelines}, though the exact same principles apply for TI.
The simplest possible alchemical pathway is a \textit{linear} pathway:

@@

However, this approach ends up being terrible for removing or adding repulsive potentials that go to infinity quickly at or near the origin. One way to look at this is to examine how low $\lambda$ must go to reduce the energy at $0.5\sigma$ (the atomic size parameter) down to 1 $k_BT$, where thermal fluctuations make it possible for other atomic sites to penetrate routinely that deep. If we are trying go from a particle being present, and desire to make it disappear alchemically, then if the repulsive terms is of the form $\epsilon\frac{\sigma}{r}^{12}$, then if $\epsilon$ was 1 $K_BT$ at the temperature of interest, and we start with the particle present, then solving for $(1-\lambda)(1 kB_T)\frac{\sigma}{\frac{1}{2}\sigma}) = 1 k_B T$ we get $\lambda = 1-2^{-12} \sim 0.999976$.  So we have gone virtually all the way to the end of the transformation, there's still an impenetrable post in the middle of our simulation! This is not very much like the desired final state of no interactions between the particle and its environment. We can play around with a few ways of modifying this, like simulating many more intermediate states near $\lambda=1$. Various analyses have shown that this is not a very good strategy, however~\cite{pham2011identifying, beutler1994avoiding, zacharias1994separationshifted, blondel2004ensemble}

@@

Where $r_{ij}$ is the distance between two particles $i$ and $j$, $\epsilon_{ij}$ and $\sigma_{ij}$ are the Lennard-Jones parameters corresponding to the interaction between particles $i$ and $j$, and $\alpha$ is a constant (0.5 is optimal for the functional form shown above).  This functional form has exactly the property we are looking for: it recovers the Lennard-Jones potential when $\lambda=1$, and the other endpoint ($\lambda=0$), it is exactly zero for all $r_{ij}$ everywhere, and as $\lambda$ goes to zero, the  $\alpha(1-\lambda)$ term lowers the infinite energy in the core.  There are several different variants of the same functional form~\cite{zacharias1994separationshifted, beutler1994avoiding,pham2011identifying}, but the one in given in eq.~\ref{eq:softcore} is easy to understand and implement and fairly numerically stable. This functional form is shown in parts C and D of figure~\ref{fig:fig_what_is_lambda}.

@@

We have discussed optimal ways of disappearing or appearing Lennard-Jones interaction sites and turning on and off electrostatics terms.  What about performing both transformations at the same time? We can't turn off the electrostatics linearly at the same time we turn off the Lennard-Jones terms, as it would leave infinitely large attractive and repulsive electrostatic terms "bare" at small $\lambda$, resulting in the simulation crashing. It \textit{is} possible to apply the same soft core approach to the Coulomb interaction, but it turns out to be somewhat difficult to turn off two soft-core terms  terms simultaneously~\cite{Steinbrecher:JCC:2011}.  We therefore recommend performing the transformations in sequence; first, turning off all electrostatics for atoms that must be removed, inserting and removing Lennard-Jones sites (both the insertion and removal can be done simultaneously), and then turning electrostatics for the introduced particles on.  However, If there are no removals or introduction to atomic sites, then it is reasonable to change all interactions linearly.  
Other issues, such as whether absolute calculations should retain or remove intramolecular nonbonded interactions, may be less critical and recommendations differ among studies in the literature~\cite{}, and reasonable efficiency can be often obtained with either choice even if some are somewhat better or worse than others. Our recommendation is to leave the intramolecular interactions on during the transformation for simplicity if there are no other known issues with this approach.  The key thing to watch out for is whether the total potential energy, and therefore the intermediate ensembles sampled, change smoothly from beginning to end.  If a configurational phase transition is noted in the process (like substantial dihedral rearrangement), then other steps must be taken, perhaps including modifying intramolecular interactions~\cite{???}.  These problems can be diagnosed by noticing lack of configuration space overlap between different simulations (see section~\ref{subsec:})" 

@@

Typically in single topology relative calculations it proves most efficient to first remove electrostatic interactions of any atoms which will be deleted, then modify other nonbonded interactions, then restore electrostatic interactions of any atoms which are being inserted.
Other schemes, such as simultaneously changing electrostatic and Lennard-Jones interactions, even with electrostatic ``soft core'' potentials~\cite{steinbrecher2007nonlinear}, in our experience typically introduce errors and/or instabilities or are at least unreliable.
It is expected that similar considerations will apply to dual topology calculations, and the principle of first removing electrostatics and then removing steric interactions will likely serve well.

@@

The spacing depends to some extent on the choice of analysis method, though states should essentially be spaced equidistant in the relevant thermodynamic length~\cite{}.

@@

@@

Absolute free energy calculations for dissimilar ligands can have particular challenges because the (potentially incorrect) assumption of consistent binding modes across a series of similar ligands is likely to be even less robust than the in the case of relatively calculations.
This means that researchers performing absolute binding free energy calculations will have to pay particular attention to generating reasonable putative binding modes.

In some cases, it is tempting to simply use docking techniques to generate initial bound structures for starting molecular dynamics simulations.
However, timescales for binding mode interconversion are usually slow compared to MD/free energy timescales, meaning that simulations started from different potential binding modes are likely to yield disparate computed binding free energies~\cite{mobley2006use, palma2012computation, mobley2012perspective, gill2018binding} .
And docking techniques are good at identifying sterically reasonable potential binding modes, but still perform relatively poorly at identifying a single dominant binding mode \emph{a priori}.~\cite{}

It is worth highlighting a recent SAMPL blind challenge on HIV integrase as an illustration of this. 
Many submissions, using state-of-the-art methods, had difficulty even predicting which \emph{binding site} ligands would bind in (most submissions placed more than half of the ligands into the incorrect binding site), and even given correct binding sites, the binding mode within each site was also quite difficult to predict~\cite{mobley2014blind}.
The best performing submission for predicting binding modes actually ended up being a human expert (aided by computational tools) with more than 10 years of experience on the particular target~\cite{voet2014combining}, rather than a fully automated approach.
While free energy calculations on this set had some success, many of the failures actually ended up being cases where the binding mode selected as input for free energy calculations was later found to be incorrect~\cite{gallicchio2014virtual}, highlighting the importance of these issues.

@@

\paragraph{Handling multiple potential binding modes within absolute calculations.}
Within absolute binding free energy calculations, multiple potential binding modes can be handled by two main strategies: Consider each binding mode separately (a separation of states strategy) or sample all binding modes within a single simulation~\cite{mobley2012perspective}.
This couples to the choice of restraints selected, as some restraints will allow transitions between binding modes and even binding sites (Section~\ref{sec:standardstate-restraints}), and others do not.

Sampling all binding modes within a single free energy calculation is usually impractical without some form of enhanced sampling or at least Hamiltonian replica exchange~\cite{wang2013identifying} because barriers for binding mode interconversion result in kinetics which are too slow compared to simulation timescales~\cite{mobley2006use, palma2012computation,mobley2012perspective, gill2018binding}.
Hamiltonian exchange, coupled with appropriate restraints, can allow the ligand to relatively rapidly exchange between potential binding modes when non-interacting, accelerating sampling of binding modes~\cite{wang2013identifying}.
\todo[inline, color={yellow!40}]{DLM probably need more background refs on Hamiltonian lambda exchange here.}

Separation of states provides a simple though potentially expensive alternative, where each stable binding mode is considered separately with a binding free energy calculation restricted to that binding mode, and then (as long as the binding modes are non-overlapping) the resulting component binding free energies can be combined into a total~\cite{mobley2006use, mobley2012perspective}.
This approach necessitates a separate binding free energy calculation for each potential binding mode, however, so it can be computationally quite costly.
If relative populations of different stable binding modes were available from some other technique, it could make this separation of states approach considerably more efficient~\cite{mobley2012perspective, gill2018binding}.

@@

To deal with this, some practitioners have actually computed relative binding free energies of different binding modes of the same ligand~\cite{palma2012computation} [ref Jorgensen (?) etc.].
For example, a mutation which adds a methyl to an aromatic ring of a larger ligand might yield one result if the methyl points in one direction, and a different value if it points in the other due to slow ring motions. [e.g. get ref from Sukanya]
One could compute the free energy of turning off the methyl group in one orientation and turning it back on in the other orientation to obtain the free energy difference between the two potential binding modes.
While this approach has precedent, it is relatively difficult to automate at present and requires considerable care.

>>>>

    \caption{\textbf{Thermodynamic cycle required for an aboslute free energy calculation -- absolute free energy of binding example} The fully interacting ligand in water (\textbf{A}), has its charges turned off to pass to \textbf{B} followed by turning of van der Waals terms, resulting in a non-interacting ligand in water in \textbf{C}. Restraints are used on the fully interacting ligand in the binding site of a protein or host molecule \textbf{D}. The next step is to turn off the charges again (\textbf{E}) followed by the van der Waals interactions resulting in a non-interacting complex state (\textbf{F}). Free energyes can be computed as $\Delta G_{bind} = (\Delta G^{\mathrm{elec}}_{\mathrm{solv}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{solv}})-(\Delta G^{\mathrm{elec}}_{\mathrm{bound}}+ \Delta G^{\mathrm{VdW}}_{\mathrm{bound}}$).

@@

For solvation free energy calculations, handling the standard state is typically straightforward, and treating it correctly simply means ensuring that the non-interacting solute is taken to the same (or equivalent) final reference state in both environments, e.g. that the transformation involves a 1 M to 1 M equivalent transfer free energy (where the non-interacting solute still occupies essentially the same volume as the solute in the interacting system).

@@

In practice, this has implications for how the calculations are done, as the reference concentration must enter the thermodynamic cycle employed, see an example of such a thermodynamic cycle in Fig.~\ref{fig:fig_absolute_thermodynamic_cycle}.

@@

In practice, a variety of types of restraints are common, from simple harmonic distance restraints between the ligand and the protein~\cite{mobley2006use}, to flat-bottom restraints which work similarly but only exert a force if the ligand leaves a specific region~\cite{chen2007can}.
Alternatively, a set of restraints proposed by Boresch have also commonly been employed, where all six rigid-body degrees of freedom governing the orientation of the ligand relative to the receptor are restrained~\cite{boresch2003absolutea, leitgeb2005alchemical}.

@@

For example, with more involved restraints, sampling at intermediate $\lambda$ values will usually not need to be as extensive but more computational effort must go to computing the restraining free energy.
Additionally, such restraints would typically keep the ligand from exploring alternative binding modes, which may be undesirable with Hamiltonian $\lambda$ exchange or expanded ensemble techniques where allowing the ligand to exchange binding modes when it is non-interacting could provide sampling benefits~\cite{wang2013identifying}.
Concretely, flat-bottom restraints might allow a ligand to explore multiple binding sites, harmonic restraints multiple binding modes within a site, and Boresch restraints a single binding mode within a single site.
See additional discussion of the possibility of multiple binding modes in Sec.~\ref{sec:multiple_binding_modes} below.

@@

If the ligand is a tight binder\todo[inline, color={red!40}]{AR: Should we give an order of magnitude to define a tight and a weak binder? ASJSM: Yes!}, it is usually safe to employ a restraint that allows some of the unbound configurations to be sampled since they generally contribute negligibly to the partition function (i.e., they have a relatively small Boltzmann weight) as long as the sampled volume is not so large that their cumulative contribution becomes significant.

@@

Many relative free energy planning tools have been set up to avoid changing the net charge of the systems considered, including LOMAP~\cite{liu2013lead} and Schr\"{o}dinger's FEP+~\cite{wang2015accurate}. Absolute free energy calculations can also potentially avoid changing the charge of the system by making a charge perturbation of equal and opposite sign elsewhere in the system; for example, as a charged ligand is removed, a charged counterion of opposite sign could also be removed, or one of the same sign could be inserted. This is sometimes referred to as an "alchemical ion" approach for dealing with the needed charge change, and is also employed by the Yank free energy package~\cite{wang2013identifying}.
Charge corrections have also been explored, and are potentially a viable solution to this problem~\cite{mey2018impact} where artifacts introduced by finite-size effects are corrected numerically.

@@

Some choices are particularly crucial -- for example, transformations involving insertions or deletions of atoms should employ a soft-core potential path for Lennard-Jones or other interactions with repulsive interactions that go to infinite energy at small radius~\cite{beutler1994avoiding, beutler1994molecular,gapsys2012new}.

@@

Another way of stating this is that intermediate states should sample molecular configurations that are as similar as possible to their neighboring states. The more similar the configurations are between intermediate states, the lower the statistical uncertainty is in the estimate of free energy between intervals. This can be proven directly from the BAR and MBAR formulas~\cite{bennett1976efficient,klimovich2015guidelines}, though the exact same principles apply for TI. For a 'good' path to work sufficient overlap between pontential energies is required. Fig.~\ref{fig:fig_what_is_lambda} \textbf{A} and \textb{B} try to illustrate this. Fig.~\ref{fig:fig_what_is_lambda} \textbf{A} shows in a pictorial way a soft-core potential can be applied across different $\lamda$s. Fig.~\ref{fig:fig_what_is_lambda} \textbf{B} illustrates the potential energy distributions at the different $\lambda$ intermediates, with sufficient overlap between neighboring $\lambda$ states to ensure that reweighting estimators such as MBAR can be used for analysis (see Sec.\ref{subsec:estimators}). The actual transformation is best handled with soft-core potentials of the form shown in Fig.\ref{fig:fig_what_is_lambda} \textbf{C} and \textbf{B}, with more details given below. 
So how do we adjust the potentials between the two end states based on $\lambda$? The simplest possible alchemical pathway is a \textit{linear} pathway:

@@

However, this approach ends up being terrible for removing or adding repulsive potentials that go to infinity quickly at or near the origin. One way to look at this is to examine how low $\lambda$ values must go to reduce the energy at $0.5\sigma$ (the atomic size parameter) down to 1 $k_BT$, where thermal fluctuations make it possible for other atomic sites to penetrate routinely that deep. If we are trying go from a particle being present, and desire to make it disappear alchemically, then if the repulsive terms is of the form $\epsilon\frac{\sigma}{r}^{12}$, then if $\epsilon$ was 1 $K_BT$ at the temperature of interest, and we start with the particle present, then solving for $(1-\lambda)(1 kB_T)\frac{\sigma}{\frac{1}{2}\sigma}) = 1 k_B T$ we get $\lambda = 1-2^{-12} \sim 0.999976$. So we have gone virtually all the way to the end of the transformation, there's still an impenetrable post in the middle of our simulation! This is not very much like the desired final state of no interactions between the particle and its environment. We can play around with a few ways of modifying this, like simulating many more intermediate states near $\lambda=1$. Various analyses have shown that this is not a very good strategy, however~\cite{pham2011identifying, beutler1994avoiding, zacharias1994separationshifted, blondel2004ensemble, gapsys2012new}.

@@

Where $r_{ij}$ is the distance between two particles $i$ and $j$, $\epsilon_{ij}$ and $\sigma_{ij}$ are the Lennard-Jones parameters corresponding to the interaction between particles $i$ and $j$, and $\alpha$ is a constant (0.5 is optimal for the functional form shown above).  This functional form has exactly the property we are looking for: it recovers the Lennard-Jones potential when $\lambda=1$, and the other endpoint ($\lambda=0$), it is exactly zero for all $r_{ij}$ everywhere, and as $\lambda$ goes to zero, the  $\alpha(1-\lambda)$ term lowers the infinite energy in the core.  There are several different variants of the same functional form~\cite{zacharias1994separationshifted, beutler1994avoiding,pham2011identifying}, but the one in given in eq.~\ref{eq:softcore} is easy to understand and implement and fairly numerically stable. This functional form is shown in parts \textbf{C} and \textbf{D} of Fig.~\ref{fig:fig_what_is_lambda}.

@@

We have discussed optimal ways of disappearing or appearing Lennard-Jones interaction sites and turning on and off electrostatics terms. What about performing both transformations at the same time? We can't turn off the electrostatics linearly at the same time we turn off the Lennard-Jones terms, as it would leave infinitely large attractive and repulsive electrostatic terms "bare" at small $\lambda$, resulting in the simulation crashing. It \textit{is} possible to apply the same soft core approach to the Coulomb interaction, but it turns out to be somewhat difficult to turn off two soft-core terms simultaneously~\cite{steinbrecher2011softcore}. We therefore recommend performing the transformations in sequence; first, turning off all electrostatics for atoms that must be removed, inserting and removing Lennard-Jones sites (both the insertion and removal can be done simultaneously), and then turning electrostatics for the introduced particles on. However, If there are no removals or introduction to atomic sites, then it is reasonable to change all interactions linearly.  
Other issues, such as whether absolute calculations should retain or remove intramolecular non-bonded interactions, may be less critical and recommendations differ among studies in the literature~\cite{}, and reasonable efficiency can be often obtained with either choice even if some are somewhat better or worse than others. Our recommendation is to leave the intramolecular interactions on during the transformation for simplicity if there are no other known issues with this approach.  The key thing to watch out for is whether the total potential energy, and therefore the intermediate ensembles sampled, change smoothly from beginning to end. If a configurational phase transition is noted in the process (like substantial dihedral rearrangement), then other steps must be taken, perhaps including modifying intramolecular interactions~\cite{???}. These problems can be diagnosed by noticing lack of configuration space overlap between different simulations (see section~\ref{sec:are-they-good})" 

@@

Typically in single topology relative calculations it proves most efficient to first remove electrostatic interactions of any atoms which will be deleted, then modify other non-bonded interactions, then restore electrostatic interactions of any atoms which are being inserted.
Other schemes, such as simultaneously changing electrostatic and Lennard-Jones interactions, even with electrostatic ``soft-core'' potentials~\cite{steinbrecher2007nonlinear}, in our experience typically introduce errors and/or instabilities or are at least unreliable.
It is expected that similar considerations will apply to dualtopology calculations, and the principle of first removing electrostatics and then removing steric interactions will likely serve well.

@@

The spacing depends to some extent on the choice of analysis method, though states should essentially be spaced equidistant in the relevant thermodynamic length~\cite{crooks2007measuringa, sivak2012thermodynamic}.

@@

\subsubsection{What different sampling schemes exist and which one should I use?}
\label{sec:sampling_schemes}
With all alchemical simulations having the requirement of sampling from different sets of $\lambda$ states in common, different approaches have been designed to achieve this. Fig.~\ref{fig:fig_sampling_scheme} illustrates the four most common schemes used. The simplest approach is defining a set of fixed replicas each representing a different $\lambda$ value that run independent simulations at each of the intermediate $\lambda$ values, see Fig.~\ref{fig:fig_sampling_scheme}(\textbf{A}). This type of scheme is currently used for AMBER TI calculations~\cite{song2019} and for Siremol/Openmm~\cite{}. A slightly more complex approach is allowing replica mixing, meaning that each independent $\lambda$ simulation is attempted to be exchanged with a neighboring $\lambda$ replica. This is based on ideas of Monte Carlo simulations of spin glasses by Swendsen and Wang~\cite{swendsen1986replica}. Using a Metropolis Hastings based acceptance criterion for exchanges means that the generated ensemble of all replicas is still sampling from the Boltzmann distribution and has been used in many different contexts for molecular simulations~\cite{sugita2000multidimensionala,sugita1999replicaexchangea, woods2003developmenta, jiang2010free}. The basic idea of the replica exchange scheme is shown in Fig.~\ref{fig:fig_sampling_scheme}(\textbf{B}). It is supported in various software packages that provide alchemical implementations, such as GROMACS~\cite{aldeghi2015accurate}, FEP+~\cite{wang2015accurate}, and NAMD~\cite{jiang2019computing}. A third approach borrows ideas from simulated tempering~\cite{marinari1992simulateda}. In this scheme a single replica rapidly explores all of $\lambda$ space by optimally working out weights that allow switching between different intermediate $\lambda$ values, as seen in Fig.~\ref{fig:fig_sampling_scheme}(\textbf{C}) and also referred to as self adjusted mixture sampling~\cite{lyubartsev1992newa, li2007simulated, tan2017optimally} and while promising, has so far only been supported in OpenMM Tools~\cite{andrearizzi2019choderalab}. The last approach available in conjunction with GROMACS makes use of non-equilibrium simulations~\cite{aldeghi2018accurate}. In this approach, only end state $\lambda$ replicas ($\lambda$=0, $\lambda=1$) are simulated at equilibrium intermediate information is generated from non-equilibrium simulations. A schematic of this approach is shown in Fig.~\ref{fig:fig_sampling_scheme} \textbf{D}.

\begin{figure}
    \includegraphics[width=0.95\columnwidth]{figures/fig8_sampl_scheme/Figure.pdf}
    \caption{\textbf{Four most common sampling strategies are available in different popular software packages.} \textbf{A}: Multiple replicas in parallel at different lambda states. \textbf{B}: Hamiltonian replica exchange scheme. \textbf{C}: Single replica scheme sampling from all $\lambda$ states. \textbf{D}: Non-equilibrium sampling scheme.}
    \label{fig:fig_sampling_scheme}
\end{figure} 
The best practice recommendation, is to use replica exchange type sampling schemes (Fig.~\ref{fig:fig_sampling_scheme}(\textbf{B})), or in absence of their availability and when looking at systems with fast mixing times parallel replica schemes are acceptable (Fig.~\ref{fig:fig_sampling_scheme}(\textbf{A})). Single replica schemes and non-equilibrium schemes are not as established yet, but are very promising.

\subsubsection{How long should I run my simulation for and what information should be saved?}
\label{sec:sim_length_information_kept}
Before launching alchemical free energy calculations it is wise to consider how convergence and completion will be assessed. Different conditions on when to stop alchemical free energy calculations should be determined, and this may require several iterative checks and therefore modifications to the calculation protocol.
One useful metric to use for termination is the expected or desired uncertainty of a desired free energy estimate, though care must be exercised should the uncertainty estimate prove unreliable.
In particular, if the rate of change in the free energy estimate is significant when this condition is met, the simulation may not be locally converged, and more sampling may be necessary to determine a stable free energy estimate which is no longer changing significantly over time. 
However, this is not the only metric which should be used, as the uncertainty only captures the information about the sampled phase space, not necessarily the entirety of the phase space.  
For example, convergence of relative free energy calculations in predictive simulations where the entire phase space is not known in advance, requires sampling the different kinetically stable states~\cite{mobley2012perspective}.  
This highlights the importance of choosing the correct thermodynamic path to ensure you sample the required thermodynamic states as discussed in Sec.~\ref{sec:important_path}.
The condition of minimizing the statistical uncertainty of different free energy estimators below a sufficient threshold should be one metric monitored over the simulation. This can be done through the uncertainty estimator built into certain analysis tools such as MBAR, or can be done though more general statistical tools like bootstrap sampling. 
A target statistical uncertainty should be chosen at the onset of the simulation to avoid excessively long simulations, or falling into the trap of running until the free energy estimate is "good enough," which is subjective and has no defined criteria. This could be a fixed value such as $0.20 \mathrm{kcal/mol}$, or a functional quantity such as "below $0.5 \mathrm{kcal/mol}$ and $10\%$ of the free energy estimate." The user does not need to monitor this information in real-time and can choose to run simulations for fixed duration (either time or number of samples) and run analysis on the data collected thus far. If more samples are needed, the simulations can be resumed, or, started again in different initial conditions. 
Convergence in other alchemical observables should also be monitored to determine if the defined phase space has been sufficiently sampled and enough decorrelated samples have been drawn. These additional observables include, but are not limited to, the variance in $\frac{dU}{d\lambda}$ across all $\lambda$ values, calculating the variance in free energy using bootstrap analysis, and comparing differences in free energies calculated using different percentages of the simulation in both the forward and reverse directions ( see Fig. \ref{fig:convergence_forward_reverse}).
    \includegraphics[width=0.95\linewidth]{figures/fig10_forward_reverse/Figure.pdf}
    \caption{\textbf{Free energy (in k$_{B}$T) for two different relative binding free energy (RBFE) perturbations.} 
    Each plot shows the estimated free energy change using a varying fraction of total simulation time (up to 5 ns total). 
    The top panel shows the three-step protocol for a perturbation involving 5 perturbed atoms, while the bottom panel shows the same protocol for a perturbation involving 10 perturbed atoms. The difference in energy between the forward (blue) and reverse (red) free energy calculations at the midpoint of the simulation time gives an indication of the overall convergence of the simulation, with differences over 1 k$_{B}$T indicating poor convergence.}
    \label{fig:convergence_forward_reverse}
Each of these metrics have demonstrated promising results for diagnosing when a simulation has a convergence issue beyond simple convergence of uncertainty estimate. 
Results obtained from calculations with convergence issues should be checked for errors or run for longer before any confidence should be placed in conclusions drawn from their analysis.
In relative calculations that share similar binding modes, for example, and do not induce large conformational changes when in complex with protein, the need to sample exhaustively to converge estimates in free energy differences is often not necessary due to the locality of sampling changes in the molecular topology and shared phase space of the core atoms.
However, even subtly induced changes in protein binding configuration will require more sampling or cause local convergence to a free energy estimate that has high error.
Many enhanced sampling techniques have been proposed to try and overcome barriers in free energy calculations and mitigate convergence to erroneous estimates~\cite{}. 
The confidence a user should have in a free energy estimate is significantly improved when both the uncertainty of the free energy estimate is low, and when other observables have reached a convergence.
Another aspect that can be monitored is bias in the simulation. A simple, but effective, way of checking for bias is to compute the same property of interest by multiple estimators (e.g. free energy difference). 
The uncertainty in the free energy, for example, has multiple ways to be estimated, e.g. through standard error propagation methods (including MBAR's estimator, which is based on the same principles as standard error propagation), through bootstrap methods, through multiple independent runs, etc. 
Independent of how the property is estimated, its important to remember that they are \textit{estimations of the property}, not the true underlying property itself. 
These estimators are usually consistent estimators, meaning they will converge to the true answer in the limit of sufficient sampling, not necessarily unbiased ones though.
As such, it is a good idea to subject different estimators to the same data to see if they yield either the same estimate (within error and bias), or if they fluctuate wildly. 
This is not a perfect method as some estimators, such as exponential averaging, will converge significantly more slowly, relative to more accurate estimators like MBAR. 
Therefore, it is a good idea to apply the estimators to different fractions of the data to see if the main estimator of free energy you have chosen is stable.
Each method requires different data from the simulation be collected.  If, for instance, the free energy estimator selected is thermodynamic integration, then values of $\frac{dU}{d\lambda}$ at uncorrelated data points must be collected. Once a combination of knowing what type of simulation you will run, which alchemical topology you will simulate, what alchemical path you will simulation along, and what your stopping conditions are, then you are ready to enumerate the information you should capture. Below is a sample of the minimal information you need for a set of common estimators (discussed in more detail in Sec.~\ref{subsec:estimators}):
    \item Thermodynamic Integration (TI) requires $\frac{\partial U(\vec{x})}{\partial\lambda}$.
    \item Exponential Averaging (EXP) needs \textit{either} $\Delta U_{k,k+1}(\vec{x})$ or $\Delta U_{k,k-1}(\vec{x})$, depending on the direction its being evaluated in.
    \item Bennett Acceptance Ratio (BAR) needs \textit{both} $\Delta U_{k,k+1}(\vec{x})$ and $\Delta U_{k,k-1}(\vec{x})$.
    \item Weighted Histogram Analysis Method (WHAM) and Multistate Bennett Acceptance Ratio (MBAR) both need the complete set of $\Delta U_{k,j} \, \forall \, j=\{1...K\}$. WHAM must have this information binned.
The potential derivative required for TI should generally be calculated during the simulation; only under every rare circumstances~\cite{naden2015linear} can it postprocessed by a code that does not evaluate the derivatives. Many codes already have options for doing this.
If that option is unavailable, you can estimate it through finite difference (if sufficient information is collected), but this  will introduce significant error, and is generally not a best practice. The BAR estimator may be a better, and simpler choice at that point as you will have at least the same level of information. 
The potential energy differences required for EXP, BAR, MBAR, and WHAM can be calculated either during the simulation or in post-processing. It is recommended to calculate the potential differences in code when possible to avoid extra overhead and possible errors produced by running the simulation twice, and to reduce the amount of stored information. 
Although TI must be usually be calculated in code, as it requires the derivative, there is one condition under which it actually has the fastest computation time. 
If the alchemical path you have chosen is a linear alchemical path, then you get $\frac{dU}{d\lambda} = U_0(\vec{x}) - U_1(\vec{x})$, which is the difference between the initial and final states. 
However, because of the problems with linear paths already discussed in this paper, this simplification is rarely that useful.

@@

Absolute free energy calculations for dissimilar ligands can have particular challenges because the (potentially incorrect) assumption of consistent binding modes across a series of similar ligands is likely to be even less robust than the in the case of relatively calculations.
This means that researchers performing absolute binding free energy calculations will have to pay particular attention to generating reasonable putative binding modes.
In some cases, it is tempting to simply use docking techniques to generate initial bound structures for starting molecular dynamics simulations.
However, timescales for binding mode interconversion are usually slow compared to MD/free energy timescales, meaning that simulations started from different potential binding modes are likely to yield disparate computed binding free energies~\cite{mobley2006use, palma2012computation, mobley2012perspective, gill2018binding} .
And docking techniques are good at identifying sterically reasonable potential binding modes, but still perform relatively poorly at identifying a single dominant binding mode \emph{a priori}.~\cite{} 
It is worth highlighting a recent SAMPL blind challenge on HIV integrase as an illustration of this. 
Many submissions, using state-of-the-art methods, had difficulty even predicting which \emph{binding site} ligands would bind in (most submissions placed more than half of the ligands into the incorrect binding site), and even given correct binding sites, the binding mode within each site was also quite difficult to predict~\cite{mobley2014blind}.
The best performing submission for predicting binding modes actually ended up being a human expert (aided by computational tools) with more than 10 years of experience on the particular target~\cite{voet2014combining}, rather than a fully automated approach.
While free energy calculations on this set had some success, many of the failures actually ended up being cases where the binding mode selected as input for free energy calculations was later found to be incorrect~\cite{gallicchio2014virtual}, highlighting the importance of these issues.

@@

\paragraph{Handling multiple potential binding modes within absolute calculations.}
Within absolute binding free energy calculations, multiple potential binding modes can be handled by two main strategies: Consider each binding mode separately (a separation of states strategy) or sample all binding modes within a single simulation~\cite{mobley2012perspective}.
This couples to the choice of restraints selected, as some restraints will allow transitions between binding modes and even binding sites (Section~\ref{sec:standardstate-restraints}), and others do not.
Sampling all binding modes within a single free energy calculation is usually impractical without some form of enhanced sampling or at least Hamiltonian replica exchange~\cite{wang2013identifying} because barriers for binding mode interconversion result in kinetics which are too slow compared to simulation timescales~\cite{mobley2006use, palma2012computation,mobley2012perspective, gill2018binding}.
Hamiltonian exchange, coupled with appropriate restraints, can allow the ligand to relatively rapidly exchange between potential binding modes when non-interacting, accelerating sampling of binding modes~\cite{wang2013identifying}.
\todo[inline, color={yellow!40}]{DLM probably need more background refs on Hamiltonian lambda exchange here.}
Separation of states provides a simple though potentially expensive alternative, where each stable binding mode is considered separately with a binding free energy calculation restricted to that binding mode, and then (as long as the binding modes are non-overlapping) the resulting component binding free energies can be combined into a total~\cite{mobley2006use, mobley2012perspective}.
This approach necessitates a separate binding free energy calculation for each potential binding mode, however, so it can be computationally quite costly.
If relative populations of different stable binding modes were available from some other technique, it could make this separation of states approach considerably more efficient~\cite{mobley2012perspective, gill2018binding}.

@@

To deal with this, some practitioners have actually computed relative binding free energies of different binding modes of the same ligand~\cite{palma2012computation}.
For example, a perturbation which adds a methyl to an aromatic ring of a larger ligand might yield one result if the methyl points in one direction, and a different value if it points in the other due to slow ring motions. [e.g. get ref from Sukanya]
One could compute the free energy of turning off the methyl group in one orientation and turning it back on in the other orientation to obtain the free energy difference between the two potential binding modes.
While this approach has precedent, it is relatively difficult to automate at present and requires considerable care.
